\papertitle

### 2013-01-24 manuscript.tex
Integrate Ethan's and some of Fernando's comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints, and typically more 
computing power. These specializations present many opportunities for cooperation. 
In this work, we demonstrate cooperative target-seeking between a 13 gram 
ornithopter MAV and a lightweight ground station using computer vision. We 
develop models for the ornithopter, ground station, and cooperative system 
dynamics. We determine model parameters of the real systems through experimental 
system identification. Finally, we verify those models using experiments on 
narrow passage traversal, and arrive at a cooperative system model which 
accurately predicts the backwards-reachable region for successfully 
negotiating ornithopther flight through narrow openings.

We introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon previous designs by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithopter by running a real-time 
motion tracking algorithm on a live video stream.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit their sensing and computational capabilities. They can 
execute only the most rudimentary control and state estimation algorithms in 
real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
Though they demonstrate safety and noise profiles superior to those of 
rotorcraft MAVs, this comes at the expense of reduced payload.

The non-linear dynamics of flapping-wing flight compound 
these constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Much 
past research on flapping-wing MAVs focuses on overcoming the difficulties 
of controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wing-beat 
period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, and using 
linear, low-dimensional models to predict flight 
behavior~\cite{Shigeoka:ornithopter}. These advances lead to the development 
of high-performance platforms such as the sub-15 gram DelFly, described by 
deCroon et al.~\cite{delfly:design}.

Many control laws previous researchers developed from dynamic models of 
flapping-wing flight, such as periodic forcing and other control methods 
described by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, 
less computationally-intensive method of control for negotiating narrow 
passages, by exploiting cooperation between specialized robotic agents.

@@

There also exists a body of work on the guidance and control of fixed- and 
rotary-wing MAV platforms, but many of these methods make extensive use of 
GPS or motion capture systems~\cite{kingston:timeattitude}\cite{kanade:3dvision}. 
Indoors, using GPS to estimate the pose of an MAV is impossible, and motion 
capture requires extensive and stationary modification of the environment. 
The power and payload limits on flapping-wing MAVs make complex on-board 
sensing difficult, so multiagent cooperative control schemes are apt 
solutions.

The literature on cooperative control of MAVs is similarly limited to larger
vehicles and purely analytical explorations. Nebot, Hyams, and Jung all 
demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Nebot2003Agents}\cite{Hyams1999Cooperative}\cite{Jung1998Range}. 
Luo and Mehta analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
Rudol demonstrated a cooperative visual servoing system similar with 
UGVs and MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of 
magnitude larger than the millirobotic systems we target, and so were able 
to take advantage of high-performance processors and specialized cameras.

The small size of millirobotic air and ground vehicles necessitates 
specialization, but it also pays dividends agility. A group of specialized 
heterogeneous millirobots can navigate more challenging terrains and 
interact with their environments in ways that larger, monolithic vehicles 
cannot. Despite this advantage, in order to collectively equal the sensing and 
computational capabilities of larger robots, they will also have to interact 
so that multiple robots can benefit from the specialized capabilities of 
cooperating vehicles. Millirobotic systems will benefit from cooperation 
even more than their larger counterparts.

@@

Rather than power-intensive PC processors, we used the ARM-based BeagleBoard 
single-board computer as a reference computation platform. The BeagleBoard 
consumes approximately one watt while running our computer vision and 
control algorithms. It uses the same processor as popular sub-10 gram 
processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a 
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
modified with a custom version of the Linux kernel. 
We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
{http://www.python.org/}} programming language for control, robot 
communication, and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
computer vision libraries for image capture and processing.
%TODO mention code is open source, available on GitHub

@@

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a 
particle filter~\cite{thrun2005probabilistic}. We initialize the particles 
uniformly across the frame before tracking begins. To improve numerical 
stability, we normalize the weights of all particles on each iteration, and 
only resample the particle population when the mean particle weight falls 
below a pre-determined threshold. 
%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across of the pixel space. This helps prevent overconfidence 
and collapse, and decreases reacquisition delay.

@@

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its 
reference attitude, which is in turn governed by its on-board attitude 
controllers (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
the estimate of 2D vehicle pose from the pose estimation algorithm, the 
ground station calculates an altitude error and horizontal error. The ground 
station then commands a reference attitude to the H$^2$Bird over radio, 
according to a PID control law. We apply a fixed-lag window the PID 
controller integral term to prevent wind-up while we launch the ornithopter. 
We tuned the proportional and derivative terms of this outer controller to 
dampen the overshoot and drift caused by the cooperative system latency.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

We determine the backwards reachable set of initial states for successful
narrow passage traversal both geometrically and using Monte Carlo simulation.
We simulate the motion of the ornithopter beginning from various positional
and angular heading initial conditions. We use our model, as determined
through system identification as described in Section~\ref{sec:system_model},
using the cooperative control method in Figure~\ref{fig:block_diagram}. The
results are presented in Figures~\ref{fig:feasible_set}
and~\ref{fig:heat_map}.

@@

Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
no successes within the geometrically infeasible region. There are a few 
successes near the boundary of the infeasible region determined in 
simulation. While there are successes along the boundary of the camera 
viewing triangle, there are also many failures. We further explore this 
behavior using Monte Carlo simulation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints, and typically more 
computing power. These specializations present many opportunities for 
robot-robot cooperation. 
In this work, we demonstrate cooperative target-seeking between a 13 gram 
ornithopter MAV and a lightweight ground station using computer vision. We 
develop models for the ornithopter, ground station, and cooperative system 
dynamics. We determine model parameters of the real systems through experimental 
system identification. Finally, we verify those models using experiments on 
narrow passage traversal, and arrive at a cooperative system model which 
accurately predicts the backwards-reachable region for successfully 
negotiating ornithopther flight through narrow passages.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon previous designs by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
control with a lightweight ground station, which has power and weight 
requirements appropriate for deployment on a ground vehicles with 10 gram 
payloads. The ground station provides heading estimates to the ornithopter by 
running a real-time motion tracking algorithm over a live video stream.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit their sensing and computational capabilities. MAVs can 
execute only the most rudimentary control and state estimation algorithms in 
real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
Though they demonstrate safety and noise profiles superior to those of 
rotorcraft MAVs, this comes at the expense of reduced payload.

The non-linear dynamics of flapping-wing flight compound 
these constraints~\cite{cheng:tandrdamping}\cite{humbert:dipteran2}. Much 
past research on flapping-wing MAVs focuses on overcoming the difficulties 
of controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wing-beat 
period~\cite{cheng:tandrdamping}\cite{Schenato:flightcontrol}, and using 
linear, low-dimensional models to predict flight 
behavior~\cite{Shigeoka:ornithopter}. These advances led to the development 
of high-performance platforms such as the sub-15 gram DelFly, described by 
deCroon et al.~\cite{delfly:design}.

Many control laws researchers have developed from dynamic models of 
flapping-wing flight, such as periodic forcing and other control methods 
described by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, 
less computationally-intensive method of control for negotiating narrow 
passages, by exploiting cooperation between specialized robotic agents.
%

@@

There also exists a body of work on the guidance and control of fixed- and 
rotary-wing MAV platforms, but many of these methods make extensive use of 
GPS or motion capture systems~\cite{kanade:3dvision}\cite{kingston:timeattitude}. 
Indoors, using GPS to estimate the pose of an MAV is unreliable, and motion 
capture requires extensive and stationary modification of the environment. 
The power and payload limits on flapping-wing MAVs make complex on-board 
sensing difficult, so multiagent cooperative control schemes are apt 
solutions.

The literature on cooperative control of MAVs is similarly limited to larger
vehicles and purely analytical explorations. Hyams et al., Jung et al. and Nebot et al. all 
demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}. 
Luo, et al. and Mehta, et al. analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
Rudol et al. demonstrated a similar cooperative visual servoing system with 
UGVs and MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of 
magnitude larger than the lightweight systems we target, and so were able 
to take advantage of high-performance processors and specialized cameras.

The small size of lightweight air and ground vehicles necessitates 
specialization, but it also pays dividends in agility. A group of specialized 
heterogeneous, lightweight robots can navigate more challenging terrains and 
interact with their environments in ways that larger, monolithic vehicles 
cannot. Despite this advantage, in order to collectively equal the sensing and 
computational capabilities of larger robots, small robots will also have to interact 
so that multiple robots can benefit from the specialized capabilities of 
cooperating vehicles.

@@

Rather than power-intensive PC processors, we used the ARM-based BeagleBoard 
single-board computer as a reference computation platform. The BeagleBoard 
consumes approximately one watt while running our computer vision and 
control algorithms. It uses the same processor as popular sub-10 gram 
processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
light enough for deployment on lightweight ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird via a 
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
modified with a custom version of the Linux kernel. 
We use the Python\footnote{\raggedright Python: Python Software Foundation, \href{http://www.python.org/}
{http://www.python.org/}} programming language for control, robot 
communication, and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
computer vision libraries for image capture and processing.
%TODO mention code is open source, available on GitHub

@@

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a 
particle filter~\cite{thrun2005probabilistic}. We initialize the particles 
uniformly across the frame before tracking begins. To improve numerical 
stability, we normalize the weights of all particles on each iteration, and 
only resample the particle population when the mean particle weight falls 
below a pre-determined threshold. 
%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across the pixel space. This helps prevent overconfidence 
and collapse, and decreases reacquisition delay.

@@

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its 
reference attitude, which is in turn governed by its on-board attitude 
controllers (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
the estimate of 2D vehicle pose from the pose estimation algorithm, the 
ground station calculates an altitude error and horizontal error. The ground 
station then commands a reference attitude to the H$^2$Bird over radio, 
according to a PID control law. We apply a fixed-lag window to the PID 
controller integral term to prevent wind-up while we launch the ornithopter. 
We tuned the proportional and derivative terms of this outer controller to 
dampen the overshoot and drift caused by the cooperative system latency.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

We determine the backwards reachable set of initial states for successful
narrow passage traversal both geometrically and using Monte Carlo simulation.
We also simulate the motion of the ornithopter beginning from various position
and angular heading initial conditions. We use our model, as determined
through system identification as described in Section~\ref{sec:system_model},
using the cooperative control method in Figure~\ref{fig:block_diagram}. The
results are presented in Figures~\ref{fig:feasible_set}
and~\ref{fig:heat_map}.

@@

Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
no successes within the geometrically infeasible region. There are few 
successes near the boundary of the infeasible region determined in 
simulation. While there are successes along the boundary of the camera 
viewing triangle, there are also many failures. We further explore this 
behavior using a Monte Carlo simulation.
